Notes on conformal manifolds

### 2025-09-29 Draft_MLFlatDirections.tex
Update the draft and add the 3d figures

	However, providing a good $\pi$ distribution may still remains an hard problem. In order to fix this point, we employ a so called Annealed Importance Sampling (AIS) \cite{neal1998annealedimportancesampling}. AIS aims at estimating $\pi$ sequentially from multiple intermediate distributions $\pi_{n}$ that smoothly transition from an initial, tractable distribution $\pi_{0}$ (chosen to be easy to sample from) to the complex target distribution $\pi$.
	These distributions are approximated by weighted samples $\{z^{(k)}_{n},w^{(k)}_{n}\}$ of particles, such that the resulting weighted average estimates $\pi_{n}$
	\begin{equation}
		\mathbb{E}_{\pi_{n}}[f] \simeq \displaystyle \sum_{k=1}^{n_{\rm particles}} \tilde{w}^{(k)}_{n}\,f(z^{(k)}_{n})
	\end{equation}
	or equivalently
	\begin{equation} \label{eq:pifromweights}
		\pi_{n}(z) \simeq \displaystyle \sum_{k=1}^{n_{\rm particles}} \tilde{w}^{(k)}_{n}\,\delta(z-z^{(k)}_{n}).
	\end{equation}
	with  $\tilde{w}^{(k)}_n  = w^{(k)}_n/\sum_{k=1}^{n_{\rm particles}} w^{(k)}_{n}$ the normalised weights. The $\pi_{n}$ are incrementally deformed using a Markov kernel to gradually reproduce the target $d$ and the particles get reweighed. This annealing process is guided by a temperature-like parameter that gradually emphasises the data likelihood, allowing for more efficient and tunable exploration of the polynomials landscape. The approximation procedure can be sum up as follow 
	\begin{equation}
	\pi_0(z) \underset{\rm Annealing}{\longrightarrow}\pi_n(z) \simeq \displaystyle \sum_{k=1}^{n_{\rm particles}} \tilde{w}^{(k)}_{n}\,\delta(z-z^{(k)}_{n}) \underset{\rm{n} \rightarrow +\infty}{\longrightarrow} \pi(z) \simeq \displaystyle \sum_{k=1}^{n_{\rm particles}} \tilde{w}^{(k)}\,\delta(z-z^{(k)}) \underset{\rm{n_{particles}} \rightarrow +\infty}{\longrightarrow} d(z)
	\nonumber
	\end{equation}
	Therefore, we start from an intial, easy to sample distribution $\pi_0$, that is parametrised by a set of particles and weights $\{z_0^{(k)},w_0^{(k)}\}$. We gradually mutate them with an annealing procedure to get a set of $\{z_n^{(k)},w_n^{(k)}\}$, which are used to parametrise $\pi_n$. As $n$ increases, the procedure converges to the distribution $\pi$, which itself is used to parametrise the target distribution $d$.

	This procedure has the disadvantage that the weights variance tends to increase, leading to weight degeneracy. We combine the previously mentionne AIS technique with SMC methods  \cite{del2006sequential,naesseth2024elementssequentialmontecarlo} to incorporate a resampling procedure, thus focussing computational effort on high-probability regions.  This combination of importance sampling, mutation and resampling maintains diversity among the particles and prevents premature convergence to suboptimal models. These features make AIS-SMC particularly well-suited for symbolic regression tasks, where the search space is not only high-dimensional but also structured and discontinuous.

>>>>

	However, providing a good $\pi$ distribution may still remains an hard problem. In order to fix this point, we employ a so called Annealed Importance Sampling (AIS) \cite{neal1998annealedimportancesampling} which is a class of Sequetial Monte Carlo algorithm (SMC) \cite{del2006sequential,naesseth2024elementssequentialmontecarlo}. AIS aims at estimating $\pi$ sequentially from multiple intermediate distributions $\pi_{n}$ that smoothly transition from an initial, tractable distribution $\pi_{0}$ (chosen to be easy to sample from) to the complex target distribution $\pi$.
	These distributions are approximated by weighted samples $\{z^{(k)}_{n},w^{(k)}_{n}\}$ of particles, such that the resulting weighted average estimates $\pi_{n}$
	\begin{equation}
		\mathbb{E}_{\pi_{n}}[f] \simeq \displaystyle \sum_{k=1}^{n_{\rm particles}} \tilde{w}^{(k)}_{n}\,f(z^{(k)}_{n})
	\end{equation}
	or equivalently
	\begin{equation} \label{eq:pifromweights}
		\pi_{n}(z) \simeq \displaystyle \sum_{k=1}^{n_{\rm particles}} \tilde{w}^{(k)}_{n}\,\delta(z-z^{(k)}_{n}).
	\end{equation}
	with  $\tilde{w}^{(k)}_n  = w^{(k)}_n/\sum_{k=1}^{n_{\rm particles}} w^{(k)}_{n}$ the normalised weights. The $\pi_{n}$ are incrementally deformed using a Markov kernel to gradually reproduce the target $d$ and the particles get reweighed. This process is guided by a temperature-like parameter that gradually emphasises the data likelihood, allowing for more efficient and tunable exploration of the polynomials landscape. The approximation procedure can be sum up as follow 
	\begin{equation}
	\pi_0(z) \underset{\rm Annealing}{\longrightarrow}\pi_n(z) \simeq \displaystyle \sum_{k=1}^{n_{\rm particles}} \tilde{w}^{(k)}_{n}\,\delta(z-z^{(k)}_{n}) \underset{\rm{n} \rightarrow +\infty}{\longrightarrow} \pi(z) \simeq \displaystyle \sum_{k=1}^{n_{\rm particles}} \tilde{w}^{(k)}\,\delta(z-z^{(k)}) \underset{\rm{n_{particles}} \rightarrow +\infty}{\longrightarrow} d(z)
	\nonumber
	\end{equation}
	Therefore, we start from an intial, easy to sample distribution $\pi_0$, that is parametrised by a set of particles and weights $\{z_0^{(k)},w_0^{(k)}\}$. We gradually mutate them with an annealing procedure to get a set of $\{z_n^{(k)},w_n^{(k)}\}$, which are used to parametrise $\pi_n$. As $n$ increases, the procedure converges to the distribution $\pi$, which itself is used to parametrise the target distribution $d$.

	This procedure has the disadvantage that the weights variance tends to increase, leading to weight degeneracy. We combine the previously mentionned AIS-SMC with a resampling procedure, thus focussing computational effort on high-probability regions.  This combination of importance sampling, mutation and resampling maintains diversity among the particles and prevents premature convergence to suboptimal models. These features make AIS-SMC particularly well-suited for symbolic regression tasks, where the search space is not only high-dimensional but also structured and discontinuous.
